Formal Verification of Real-Time Data Processing of the LHC Beam Loss Monitoring System: A Case Study

### 2011-03-24 fmics/paper.tex
revise the last changes

Our goal is to build a generic model of the SRS structure, and to prove that it satisfies its specification, i.e. it calculates approximate running sums of received count values within acceptable error margins.
%Any statement in formal logic is necessarily precise, and one of our challenges has been characterizing our goal precisely.
%One of our challenges has been characterizing this goal precisely
Let $\mathsf{output} \, n$ denote the sum of the last $m$ received count values and $\mathsf{RS} \, n$ denote one of the outputs computed by the slices in the structure of SRS which is supposed to compute the sum of the last $m$ received count values with some delay.
A sketch of the desired correctness statement is \[\forall{n} \cdot \, \mathsf{RS} \,n= \mathsf{output} \, n \pm\text{acceptable error}\]
Although Figure~\ref{fig:srs} suggests that in the SRS structure there are only twelve outputs  (i.e. $1\leq{n}\leq11$), we would obtain a more generic result (that is useful in future upgrades of the system) if we prove the correctness of the above statement for all values of $n$.

%To prove the equation above, we need to formalize all of its terms.
In order to make the above correctness statement more precise, we need to include the notion of time.
The count values arrive at the input of the SRS block every 40 micro-seconds that is abstracted by one time step in our logical model.
We formalize the input stream to SRS as a function of time: $D\;t$ denotes the input value to the SRS structure at time $t$.
In addition, the terms in the above statement depend on the stream of input counts over time.
With these refinements, the correctness statement is defines as: \[\forall{D\,n\,t} \cdot \,\mathsf{RS} \,D\,n\,t = \mathsf{output} \,D\,\,n\, t \pm\text{acceptable error}\]

@@

%To formalize the term on the left hand side, we need to build a logical model of the SRS structure.
The first natural step to prove the correctness of the above statement is to build a logical model of the SRS structure.
Our work proceeded in two phases, starting with a model of a simplified structure and proceeding to a model which is closer to the SRS implementation in BLMS.
Our main approach in building the logical model of SRS (in both phases) was to model each building block of the SRS structure that holds a count value--for example, each cell in each shift register--as a function in HOL that takes the input stream and current time step as inputs and returns the value which the corresponding building block holds at that time.

@@

Our simplifications in the first phase include:
\begin{enumerate}
\item Masking the middle taps in each slice,
\item Considering a more regular arrangement of slices than the SRS implementation in BLMS,
\item Masking the subtractors and the accumulators in each slice, instead, defining the output of a slice as the sum of the values in its cells directly,
\item Simplifying timing by ignoring some of the delays that each building block may introduce,
\item Using natural numbers, rather than finite words, for count values
\end{enumerate}

Building a logical model for simplified structure of the SRS helped us improve our understanding of the behavior of the SRS.
{\bf what else did we learn through this process? If there was no particular advantage, I don't think we should mention the simplified version}.

In the second phase, we eliminated the first four simplifications (listed above) by adding middle taps, a more flexible arrangement of slices, the subtractor and accumulator, and considering the delays that each block may introduce.
In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure of SRS in BLMS in the following aspects: we model an infinite number of slices (rather than six slices), each with an infinite number of shift register cells and taps (rather than fixed width shift registers and only two taps), simply by letting indices range over the natural numbers without explicitly giving limits. Such a parameterization would make the model and the analysis applicable to the future upgraded/modified versions of the system.

>>>>

Our goal is to build a generic model of the SRS structure, and to prove that it satisfies its specification, that it calculates approximate running sums of received count values within acceptable error margins.
%Any statement in formal logic is necessarily precise, and one of our challenges has been characterizing our goal precisely.
%One of our challenges has been characterizing this goal precisely
Let $\mathsf{RS}\;n$ denote, as in Figure~\ref{fig:srs}, the output of slice $n$ in the SRS structure, which is supposed to compute a sum of received count values with some delay, and let $\mathsf{exact}\;n$ denote this sum.
A sketch of the desired correctness statement is \[\forall{n} \cdot \, \mathsf{RS}\;n= \mathsf{exact}\; n \pm\text{acceptable error}\]
Although Figure~\ref{fig:srs} suggests that the SRS structure has only twelve outputs  (i.e. $0\leq{n}\leq11$), we obtain a more generic result (that is useful in future upgrades of the system) by proving the correctness of the above statement for all values of $n$.

%To prove the equation above, we need to formalize all of its terms.
To make the above correctness statement more precise, we need to include the notion of time.
The count values arrive at the input of the SRS block every 40 micro-seconds, which we abstract as a single time step in our logical model.
We formalize the input stream as a function of time: $D\;t$ denotes the input value to the SRS structure at time $t$.
In addition, the terms in the above statement depend on this stream of input counts.
With these refinements, the correctness statement is: \[\forall{D\,n\,t} \cdot \,\mathsf{RS}\;D\;n\;t = \mathsf{exact}\;D\;\;n\;t\pm\text{acceptable error}\]

@@

%To formalize the term on the left hand side, we need to build a logical model of the SRS structure.
The first natural step to prove the correctness of the above statement is to build a logical model of the SRS structure.
Our work proceeded in two phases, starting with a model of a simplified structure and proceeding to a model which is closer to the SRS implementation in BLMS.
In both phases, we model each building block of the SRS structure that holds a count value--for example, each cell in each shift register--as a function in HOL that takes the input stream and current time step as inputs and returns the value the corresponding building block holds at that time.

@@

Our simplifications in the first phase include:
\begin{enumerate}
\item Masking the middle taps in each slice,
\item Considering a more regular arrangement of slices than the SRS implementation in BLMS,
\item Masking the subtractors and the accumulators in each slice, instead, defining the output of a slice as the sum of the values in its cells directly,
\item Simplifying timing by ignoring some of the delays that each building block may introduce, {\bf this is not true - there were delays in the simplified model}
\item Using natural numbers, rather than finite words, for count values
\end{enumerate}

Building a logical model for simplified structure of the SRS helped us improve our understanding of the behavior of the SRS.
{\bf what else did we learn through this process? If there was no particular advantage, I don't think we should mention the simplified version.
(I don't think there was any particular advantage -- Ramana)}.

In the second phase, we eliminated the first four simplifications (listed above) by adding middle taps, a more flexible arrangement of slices, the subtractor and accumulator, and considering the delays that each block may introduce.
In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure of SRS in BLMS in the following way: we model an infinite number of slices (rather than six slices), each with an infinite number of shift register cells and taps (rather than fixed width shift registers and only two taps), simply by letting indices range over the natural numbers without explicitly giving limits.
Such a parameterization would make the model and the analysis applicable to the future upgraded/modified versions of the system.
